---
title: "EDS 230 Lab 2 (Template)"
format: html
editor_options: 
  chunk_output_type: console
---

# **Goal:** visualize differences in male vs. female salaries across occupations

## Step 0. Load required packages & read in data
```{r}
#..........................load packages.........................
library(tidyverse)

#......................read in cleaned data......................
jobs <- readRDS(here::here("data", "jobs.rds"))
```

## **Step 1:** What data type(s)?

`occupation` is categorical
`total_earnings_male` & `total_earnings_female` ($ earned by males and females) are numerical

## **Step 2:** What visualization type (using https://www.data-to-viz.com/)?

dotplot

## **Step 3:** Create a basic plot

```{r}
wages_plot <- ggplot(jobs) +
  
  # create dumbbells ----
  geom_point(aes(x = total_earnings_female, y = occupation)) +
  geom_point(aes(x = total_earnings_male, y = occupation)) +
  geom_segment(aes(x = total_earnings_female, xend = total_earnings_male,
                   y = occupation, yend = occupation))

wages_plot
```

## **Step 4:** Improve clarity and take-home message(s)

Some improvements to make: 
- color-code male vs. female points
- reorder occupations (e.g. arrange them in descending order of average salary)
- facet wrap by our three groupings (`perc_group:` occupations that are 75%+ female, 45-55% female, and 75%+ male)
- reformat x-axis values (wages in dollars, $)

```{r}
wages_plot <- ggplot(jobs) +
  
  # create dumbbells ----
  geom_point(aes(x = total_earnings_female, y = occupation)) +
  geom_point(aes(x = total_earnings_male, y = occupation)) +
  geom_segment(aes(x = total_earnings_female, xend = total_earnings_male,
                   y = occupation, yend = occupation))

wages_plot
```

## **Step 5:** Final touches/polishing

- remove unnecessary axis labels
- add a title, subtitle, caption
- replace legend with colored-coded text 

```{r}

```

































```{r}
wages_plot <- ggplot(jobs) +
  
  # create dumbbells ----
  geom_segment(aes(x = total_earnings_female, xend = total_earnings_male,
                   y = reorder(occupation, avg_salary), yend = occupation)) +
  geom_point(aes(x = total_earnings_female, y = occupation), color = "purple", size = 2.5) +
  geom_point(aes(x = total_earnings_male, y = occupation), color = "orange", size = 2.5) +
  
  # facet_wrap by occupation grouping ----
  facet_wrap(~perc_group_label, nrow = 3, scales = "free_y") + 
  
  # axis breaks & labels -----
  scale_x_continuous(labels = scales::label_dollar(scale = 0.001, suffix = "k"),
                     breaks = c(25000, 50000, 75000, 100000, 125000)) + 
  
    # title & subtitle (add styling) ----
  labs(title = "**Earnings by Occupation and Sex**",
       subtitle = "Median earnings of full-time <span style='color:orange;'>**male**</span> versus <span style='color:purple;'>**female**</span> workers by occupation in 2016",
       caption = "<span style='color:#919092;'>*Data courtesy of TidyTuesday (March 5, 2019)*</span>") +
  
  # set theme ----
  theme_minimal() +
  
  # customize theme ----
  theme(
  
    # facet panels 
    strip.text.x = element_text(face = "bold"),
  
    # title & subtitle 
    plot.title.position = "plot", # left-align plot title with left-edge of plot, not y-axis (see: https://github.com/tidyverse/ggplot2/issues/3252)
    plot.title = ggtext::element_textbox_simple(size = 20, padding = margin(t = 0, r = 0, b = 10, l = 0)), # enables markdown text & creates textbox
    plot.subtitle = ggtext::element_textbox_simple(size = 13, padding = margin(t = 0, r = 0, b = 10, l = 0)), # enable markdown text & creates textbox
    plot.caption = ggtext::element_markdown(margin = margin(t = 15, r = 0, b = 0, l = 0)), # enable markdown text 
  
    # axes 
    axis.title = element_blank()
  
  )

wages_plot
```

